---
title: "Data Wrangling"
date: 9/18/18
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

library(tidyverse)
```

## Import FAS csv files

Import csv file and format names

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```

## Sept. 20 Lecture



## Example

Loading in new data set

```{r}
pups_data = read_csv(file = "./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```

## Look at Data

Look at litters data

```{r}
head(litters_data)
tail(litters_data)

skimr::skim(litters_data)
```

View function allows you to view entire data set in console, could cause errors in code chunks.

```{r}
head(pups_data)
tail(pups_data)

skimr::skim(pups_data)
```


## Other File Formats

Read in MLB data: Excel

```{r}
mlb_data = readxl::read_excel(path = "./data/mlb11.xlsx")

mlb_subset = readxl::read_excel(path = "./data/mlb11.xlsx", range = "A1:E17")
```

Read in Pulse Data: SAS 

```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat")
```

## Sept. 20, 2018 Lecture

```{r}
litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

```

## Select variables

By listing:
```{r, eval=FALSE}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, litter_number, gd0_weight, pups_survive)

```

By specifying a range:

```{r, eval=FALSE}
select(litters_data, gd_of_birth:pups_survive)
select(litters_data, group, gd_of_birth:pups_survive)

```

By saying what you want to remove

```{r, eval=FALSE}
select(litters_data, -litter_number)
```

Selecting and renaming simultaneously or separately:

```{r, eval=FALSE}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, gest_day_0_weight = gd0_weight)
```

Using select helpers:

```{r, eval=FALSE}
select(litters_data, starts_with("gd"))

select(litters_data, litter_number, starts_with("gd"))

select(litters_data, litter_number, starts_with("pup"))

select(litters_data, litter_number, everything())

```


## Filter Observations

```{r, eval=FALSE}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, is.na(gd0_weight))
filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low8")
filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)
filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))

filter(litters_data, sex == 1)
filter(litters_data, sex == 2, pd_walk < 11)

```

## Mutate

Creating variables and modifying existing variables:

```{r, eval=FALSE}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group))

mutate(litters_data,
       wt_gain = gd18_weight = gd0_weight,
       wt_gain_squared = wt_gain^2
       )
```

## Arrange

Arranging the data:

```{r, eval=FALSE}
arrange(litters_data, gd0_weight)
arrange(litters_data, pups_born_alive, gd0_weight)
arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

## Piping

Looking at intermediate object approach:

```{r, eval=FALSE}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Using nested function calls:

```{r, eval=FALSE}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Using piping: shift+command+m

```{r, eval=FALSE}
read_csv(".data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight = gd0_weight, 
    group = tolower(group)
  ) %>% 
  filter(!is.na(gd0_weight))
```

Illustrate placeholder use with `lm`

```{r, eval=FALSE}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight, 
    group = tolower(group)
  ) %>% 
  filter(!is.na(gd0_weight)) %>% 
  lm(gd18_weight - gd0_weight, data = .)

```






